---
title: "DE_beta2"
output: html_document
date: "2025-04-15"
---

```{r}

library(Seurat)
library(tidyverse)
library(patchwork)
library(tidyverse)
library(ggplot2)
library(SeuratData)
InstallData("panc8")


# Get system info to determine data path
sys_info <- Sys.info()

if (grepl("Linux", sys_info["sysname"])) {
  data_dir <- '../../../data/'
} else if (sys_info["sysname"] == "Windows") {
  data_dir <- 'C:/Users/hostp/Desktop/data/'
}

# Path to saved Seurat object
# We are using a subsample of the cao dataset where the 
# dublicate rownames i.e genes has been merged/aggregated
seurat_path <- file.path(data_dir, "cao_sub_agg.rds")

# Check if the object already exists
if (file.exists(seurat_path)) {
  message("Loading existing Seurat object...")
  cao_subsample <- readRDS(seurat_path) 
} else {
  print("Run cao exploration, to create seurat object")
}

panc <- LoadData('panc8')
```

### From the following a seurat object dataset consisting of both a subsample from the cao dataset and the pancreas only dataset is integrated into the same dataset

```{r}
# STEP 2: Tag datasets for tracking
panc$dataset <- "pancreas"
cao_subsample_agg$dataset <- "multi_organ"

# Preserve Organ info
panc$Organ <- "pancreas"  # static
cao_subsample_agg$Organ <- cao_subsample@meta.data$Organ[match(colnames(cao_subsample_agg), colnames(cao_subsample))]

# Preserve cell type info from pancreas dataset
panc$celltype_panc <- panc$celltype  # rename for clarity

# STEP 3: Normalize and find variable features
panc <- NormalizeData(panc)
panc <- FindVariableFeatures(panc)

cao_subsample_agg <- NormalizeData(cao_subsample_agg)
cao_subsample_agg <- FindVariableFeatures(cao_subsample_agg)

# STEP 4: Integration
features <- SelectIntegrationFeatures(object.list = list(cao_subsample_agg, panc))
anchors <- FindIntegrationAnchors(object.list = list(cao_subsample_agg, panc), anchor.features = features)
combined <- IntegrateData(anchorset = anchors)

# STEP 5: Downstream analysis
DefaultAssay(combined) <- "integrated"

combined <- ScaleData(combined)
combined <- RunPCA(combined)
combined <- RunUMAP(combined, dims = 1:30)
combined <- FindNeighbors(combined, dims = 1:30)
combined <- FindClusters(combined, resolution = 0.5)

# STEP 6: Metadata check
print(table(combined$dataset))
print(table(combined$Organ))
print(table(combined$celltype_panc, useNA = "ifany"))

# Optional: Save the integrated object
saveRDS(combined, file.path(data_dir, "combined_integrated.rds"))

```

### Differential gene expression between betacells from the panc dataset vs panc cell from cao dataset

```{r}



# Beta cells from pancreas-only dataset
beta_panc <- subset(combined, subset = celltype == "beta" & dataset == "pancreas")

# All pancreas cells from multi-organ dataset (cao)
pancreas_cao <- subset(combined, subset = Organ == "Pancreas" & dataset == "multi_organ")




```

From chat

🧬 3. Merge them into one object
r
Copy
Edit
# Combine into one object
comparison_obj <- merge(beta_panc, y = pancreas_cao)

# Create a label for comparison
comparison_obj$group <- paste0(comparison_obj$dataset, "_", comparison_obj$celltype)
Idents(comparison_obj) <- "group"

# See identity levels
levels(Idents(comparison_obj))
You should now see identities like:

csharp
Copy
Edit
[1] "pancreas_beta"      "cao_alpha" "cao_delta" ...
🔬 4. Run DE between beta_panc and all cao pancreas cells
You can either:

Keep the subtypes (alpha, delta, etc.) and compare all of them together

Or assign them a shared identity so they're treated as one group

Option A — Compare beta vs. all cao pancreas
r
Copy
Edit
# Assign new labels: beta_panc vs multi_pancreas
comparison_obj$group2 <- ifelse(comparison_obj$dataset == "pancreas", "beta_panc", "multi_pancreas")
Idents(comparison_obj) <- "group2"

# Run differential expression
de_results <- FindMarkers(comparison_obj, ident.1 = "beta_panc", ident.2 = "multi_pancreas")
head(de_results)
📊 Optional: Plot some DEGs
r
Copy
Edit
# Top DE gene
top_gene <- rownames(de_results)[1]

VlnPlot(comparison_obj, features = top_gene, group.by = "group2", pt.size = 0.1)
Let me know if you’d like to:

Export DE results to CSV

Plot more genes

Use only variable features or set thresholds

Or check for batch effects

I'm happy to help refine it based on what you want to explore 🧪📈



