---
title: "DE_beta2"
output: html_document
date: "2025-04-15"
---

```{r}

library(Seurat)
library(tidyverse)
library(patchwork)
library(tidyverse)
library(ggplot2)
library(SeuratData)
InstallData("panc8")


# Get system info to determine data path
sys_info <- Sys.info()

if (grepl("Linux", sys_info["sysname"])) {
  data_dir <- '../../../data/'
} else if (sys_info["sysname"] == "Windows") {
  data_dir <- 'C:/Users/hostp/Desktop/data/'
}

# Path to saved Seurat object
# We are using a subsample of the cao dataset where the 
# dublicate rownames i.e genes has been merged/aggregated
seurat_path <- file.path(data_dir, "cao_sub_agg.rds")

# Check if the object already exists
if (file.exists(seurat_path)) {
  message("Loading existing Seurat object...")
  cao_subsample <- readRDS(seurat_path) 
} else {
  print("Run cao exploration, to create seurat object")
}

panc <- LoadData('panc8')
cao_subsample_agg <- cao_subsample
```

### From the following a seurat object dataset consisting of both a subsample from the cao dataset and the pancreas only dataset is integrated into the same dataset

```{r}
# STEP 2: Tag datasets for tracking
panc$dataset <- "pancreas"
cao_subsample_agg$dataset <- "multi_organ"

# Preserve Organ info
panc$Organ <- "pancreas"  # static
cao_subsample_agg$Organ <- cao_subsample@meta.data$Organ[match(colnames(cao_subsample_agg), colnames(cao_subsample))]

# Preserve cell type info from pancreas dataset
panc$celltype_panc <- panc$celltype  # rename for clarity

# STEP 3: Normalize and find variable features
panc <- NormalizeData(panc)
panc <- FindVariableFeatures(panc)

cao_subsample_agg <- NormalizeData(cao_subsample_agg)
cao_subsample_agg <- FindVariableFeatures(cao_subsample_agg)

# STEP 4: Integration
features <- SelectIntegrationFeatures(object.list = list(cao_subsample_agg, panc))
anchors <- FindIntegrationAnchors(object.list = list(cao_subsample_agg, panc), anchor.features = features)
combined <- IntegrateData(anchorset = anchors)

# STEP 5: Downstream analysis
DefaultAssay(combined) <- "integrated"

combined <- ScaleData(combined)
combined <- RunPCA(combined)
combined <- RunUMAP(combined, dims = 1:30)
combined <- FindNeighbors(combined, dims = 1:30)
combined <- FindClusters(combined, resolution = 0.5)

# STEP 6: Metadata check
print(table(combined$dataset))
print(table(combined$Organ))
print(table(combined$celltype_panc, useNA = "ifany"))

# Optional: Save the integrated object
saveRDS(combined, file.path(data_dir, "combined_integrated.rds"))

```

### Top 10 differentially expressed gene between beta cells from the pancreas only dataset and the pancreas cells from the multi organ dataset

-   **IAPP (Islet Amyloid Polypeptide)** – Often co-expressed with insulin in pancreatic beta cells; plays a role in glucose regulation and satiety. Strong marker for **beta cells**.

-   **DLK1 (Delta Like Non-Canonical Notch Ligand 1)** – Involved in development and stem cell regulation; has roles in **pancreatic endocrine differentiation**.

-   **ADCYAP1 (Adenylate Cyclase Activating Polypeptide 1)** – Neuropeptide with anti-inflammatory and neuroprotective roles; also expressed in **pancreatic islets**.

-   **HAPLN4 (Hyaluronan and Proteoglycan Link Protein 4)** – Part of the extracellular matrix; not well studied in pancreas, but possibly related to structural or signaling roles.

-   **KCNG3 (Potassium Voltage-Gated Channel Modifier Subfamily G Member 3)** – Regulates electrical excitability, could be relevant in **secretory cells** like beta cells.

-   **CDKN1C (Cyclin Dependent Kinase Inhibitor 1C)** – Also known as **p57**; important in **cell cycle arrest and beta cell differentiation**. Commonly used beta cell marker.

-   **HHATL (Hedgehog Acyltransferase-Like)** – Possibly involved in signaling pathways, especially during development.

-   **SFTPA1 (Surfactant Protein A1)** – Typically associated with the **lung**, but expression in the pancreas may indicate a stress or injury response, or contamination.

-   **CAPN13 (Calpain 13)** – A protease, not well characterized in the pancreas; might play a role in **intracellular signaling or cytoskeletal remodeling**.

-   **OXGR1 (Oxoglutarate Receptor 1)** – A G protein-coupled receptor; less is known about its pancreatic role, could be involved in metabolic sensing.

```{r}
# Beta cells from pancreas-only dataset
beta_panc <- subset(combined, subset = celltype == "beta" & dataset == "pancreas")

# All pancreas cells from multi-organ dataset (cao)
pancreas_cao <- subset(combined, subset = Organ == "Pancreas" & dataset == "multi_organ")


# Keep only intersecting genes
common_genes <- intersect(rownames(beta_panc), rownames(pancreas_cao))

beta_panc <- subset(beta_panc, features = common_genes)
pancreas_cao <- subset(pancreas_cao, features = common_genes)

# Now try merging
comparison_obj <- merge(beta_panc, y = pancreas_cao)
# Merge the two Seurat objects
comparison_obj <- merge(beta_panc, y = pancreas_cao)

# Create a new group label combining dataset and cell type
comparison_obj$group <- paste0(comparison_obj$dataset, "_", comparison_obj$celltype)
Idents(comparison_obj) <- "group"

# Check identity levels
levels(Idents(comparison_obj))


# Create a simplified group: beta vs all other pancreas cells
comparison_obj$group2 <- ifelse(comparison_obj$dataset == "pancreas", "beta_panc", "multi_pancreas")
Idents(comparison_obj) <- "group2"

# Run differential expression analysis
de_results <- FindMarkers(comparison_obj, ident.1 = "beta_panc", ident.2 = "multi_pancreas")

# Get the top DE gene
top_gene <- rownames(de_results)[1]

# Plot expression across the two groups
VlnPlot(comparison_obj, features = top_gene, group.by = "group2", pt.size = 0.1)

de_results %>%
  filter(avg_log2FC > 0) %>%
  arrange(desc(avg_log2FC)) %>%
  rownames() %>% 
  head(10)

```

